### 2020-01-29 README.Rmd
add README

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Casting multiple shadows: interactive data visualisation with tours and embeddings

`liminal` aims to increase the effective use of embeddings by combining
them with interactive and dynamic graphics (via the vegawidget and shiny
packages) and with a technique from multivariate stastics called the tour (via
the tourr package).  Briefly, a tour is a sequence of interploated projections
of multivariate data onto lower dimensional space. The sequence is displayed as
a dynamic visualisation, and enables us to see the shadows the high dimensional
data makes in a lower dimensional view. By combining the tour with embedding
algorithms, we can see the following:

1. whether distances in the embedding view are meaningful 
2. the local and global structure of the data 
3. identify 'interesting' shapes or points in the data

## Outline

- review of embedding algorithms, interactive graphics, tours
- technology
- design
- examples/benchmarks
  - Kevin moon
  - pdfsense
  - single cell
  - quick draw