### 2020-01-29 Rmd/paper.Rmd
paper outline

---
title: " Casting multiple shadows: interactive data visualisation with tours and embeddings"
description: |
  A new article created using the Distill format.
author:
  - name: Stuart Lee
    affiliation: Monash University
  - name: Di Cook
    affiliation: Monash University
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```