### 2020-01-30 Rmd/paper.Rmd
outline in place

---
title: " Casting multiple shadows: interactive data visualisation with tours and embeddings"
description: |
  A new article created using the Distill format.
author:
  - name: Stuart Lee
    affiliation: Monash University
  - name: Di Cook
    affiliation: Monash University
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

<!--
- review of embedding algorithms, interactive graphics, tours
- technology
- design
- examples/benchmarks
  - Kevin moon
  - pdfsense
  - single cell
  - quick draw
-->

# Introduction

## Goals of high dimensional vis

## How do analysts actually use these methods

# Related Work

## Embedding methods

## Tours

## Interactive graphics

# Methods

## Tours as a streaming data problem

## Multiple views, multiple contexts

## Interactions and visual analytics for embeddings

# Results

## Case Study 1: Fake trees

## Case Study 2: non-spherical clusters, pdfsense

## Case Study 3: well-clustered data, single cell

## Case Study 4: quickdraw